Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-04-27 ThesisJP/Introduction/introduction.tex
Feedback from Jo Inge finished.

\section{Motivation}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal super computers, comprised in a single graphics card (\nom{GPU}{Graphics processing unit}). The latest graphics cards\footnote{As of January 2014.} from both \nom{Nvidia}{Graphics card producer} and \nom{AMD}{Graphics card producer} provide more than 5 T\nom{FLOPS}{Floating point operations per second} single precision and 1.5 TFLOPS double precision computations. For double precision, this is more than the world's most powerful supercomputer could provide in 1998, and that was a room full of computers.

@@

The estimation of the spatial covariance matrix receives special attention in this study. In previous literature on Capon beamforming the matrix inversion has always been regarded as the most computationally complex step. In this paper we show that estimation of the sample covariance matrix is the most complex part when spatial and temporal smoothing is included. We then show how the computational complexity of this estimation procedure can be reduced from cubic to square. Finally, an in-depth analysis of the arithmetic throughput on multiple platforms is given. Despite our efforts, the number of effective FLOPS is only 4 \% and 14 \% of the theoretical throughput of the target GPU for the matrix equation solver and covariance matrix estimator respectively. Still, the reported throughput of 1 Mpx/s on a high-end GPU is enough to provide real-time processing for sonar imaging.

>>>>

\section{Motivation}
A rapid development in computer game technology and accompanying programming languages have recently provided researchers with small personal super computers, comprised in a single graphics card (\nom{GPU}{Graphics processing unit}). The latest graphics cards\footnote{As of January 2014.} from both \nom{Nvidia}{Graphics card vendor} and \nom{AMD}{Graphics card vendor} provide more than 5 T\nom{FLOPS}{Floating point operations per second} single precision and 1.5 TFLOPS double precision computations. For double precision, this is more than the world's most powerful supercomputer could provide in 1998, and that was a room full of computers.

@@

The estimation of the spatial covariance matrix receives special attention in this study. In previous literature on Capon beamforming the matrix inversion has always been regarded as the most computationally complex step. In this paper we show that estimation of the sample covariance matrix is the most complex part when spatial and temporal smoothing is performed with common parameters. We then show how the computational complexity of this estimation procedure can be reduced from cubic to square. Finally, an in-depth analysis of the arithmetic throughput on multiple platforms is given. Despite our efforts, the number of effective FLOPS is only 4 \% and 14 \% of the theoretical throughput of the target GPU for the matrix equation solver and covariance matrix estimator respectively. Still, the reported throughput of 1 Mpx/s on a high-end GPU is enough to provide real-time processing for sonar imaging.
